### 2024-03-27 code/analysis-code/statistical-analysis.qmd
project

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

```{r}
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(rsample)

```

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)

```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
set.seed(123)
data_split <- initial_split(madaproject, prop = 0.7)
data_train <- training(data_split)
data_test <- testing(data_split)
```

```{r}
set.seed(123)
folds <- vfold_cv(data_train, strata = tb)
```

I will define my recipe using a recipe package under tidy models
```{r}
rec <- logistic_reg()
tb_recipe <- recipe(tb ~ ., data = data_train) %>%
  step_dummy(all_nominal_predictors())

```

```{r}
tb_workflow <- workflow(tb_recipe, log_model)
```

defining our logistic model
```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
samplefit <- fit_resamples(tb_workflow, folds, control = ctrl_preds)

```

```{r}
#augment(samplefit) %>%
#  roc_curve(tb, .pred_TB)  %>%
#  autoplot()
```

### 2024-03-27 code/eda-code/eda.qmd
project

---
title: "Predicting of the pulmunary TB cases using clinical"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

# Setup

```{r}
# Load required libraries
library(tidymodels)
library(readxl)
library(gtsummary)
library(ggplot2)
library(forcats)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
madaproject <- readRDS(data_location)
colnames(madaproject)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(madaproject)
print(summary_df)
# save to file
summarytable_file = here("results", "tables","summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Data exploration through figures

Histogram plots for the continuous outcomes.

I will look at age

```{r}
# Create the ggplot age plot
ageplot <- madaproject %>%
  ggplot(aes(x = Age, fill = ..count..)) +
  geom_histogram(color = "black", bins = 30) +  # Adjust 'bins' as needed
  scale_fill_gradient(low = "blue", high = "red")  # Specify gradient colors
# Plot the age distribution histogram
print(ageplot)

# Specify the file path to save the plot
figure_file <- here("results", "figures", "age_distribution.png")

# Save the plot as a PNG file
ggsave(filename = figure_file, plot = ageplot)

```

Now Oxygen consumption

```{r}
oxygencons <- madaproject %>% ggplot(aes(x=oxygencons, fill = ..count..)) + geom_histogram() 
plot(oxygencons)
figure_file = here("results","figures", "oxygenconsumption.png")
ggsave(filename = figure_file, plot=oxygencons) 
```

```{r}
# Boxplot for Age vs tb
age_boxplot <- ggplot(madaproject, aes(x = tb, y = Age, fill = tb)) +
  geom_boxplot() +
  labs(title = "Boxplot of Age vs. tb",
       x = "tb",
       y = "Age") +
  scale_fill_discrete(name = "tb")
figure_file = here("results","figures", "age_boxplot.png")
ggsave(filename = figure_file, plot=age_boxplot) 
```

```{r}

# Boxplot for oxygencons vs tb
oxygencons_boxplot <- ggplot(madaproject, aes(x = tb, y = oxygencons, fill = tb)) +
  geom_boxplot() +
  labs(title = "Boxplot of oxygencons vs. tb",
       x = "tb",
       y = "oxygencons") +
  scale_fill_discrete(name = "tb")
figure_file = here("results","figures", "oxygencons_boxplot.png")
ggsave(filename = figure_file, plot=oxygencons_boxplot) 
```

```{r}
library(ggplot2)

# Create scatter plot
scatter_plot <- ggplot(madaproject, aes(x = oxygencons, y = Age, color = tb)) +
  geom_point() +
  labs(title = "Age vs. Oxygen Consumption Stratified by TB Status",
       x = "Oxygen Consumption",
       y = "Age",
       color = "TB Status")

figure_file = here("results","figures", "age_oxygenconsscatter_plot.png")
ggsave(filename = figure_file, plot=scatter_plot) 
```

```{r}

# Create gtsummary table 
table1 <- 
  madaproject %>%
  tbl_summary(
    include = all_of(c("Gender", "Asthma", "Smoke", "Alcohol", "feverstatus", 
                       "Weightloss", "coughstatus", "Sputumproduction", 
                       "BloodinSputum", "cxrconclusion_chestxray", 
                       "homefuel", "hivstatus", "tb")),
    missing = "no",
    label = list(
      Asthma ~ "Asthma",
      Smoke ~ "Smoking Status",
      Alcohol ~ "Alcohol",
      feverstatus ~ "Fever Status",
      Weightloss ~ "Weight Loss",
      coughstatus ~ "Cough Status",
      `Sputumproduction` ~ "Sputum Production",
      `BloodinSputum` ~ "Blood in Sputum",
      cxrconclusion_chestxray ~ "CXR Conclusion",
      homefuel ~ "Home Fuel",
      hivstatus ~ "HIV Status",
      tb ~ "TB"
    )
  ) %>%
  add_n()

# Print the table
table1

# Save table
saveRDS(table1, file = here::here("results", "tables", "table1.rds"))

```
This is just my code to use a for loop 
```{r}

# Define variable types
categorical_vars <- c("Gender", "Asthma", "Smoke", "Alcohol", "feverstatus", 
                      "Weightloss", "coughstatus", "Sputumproduction", 
                      "BloodinSputum", "cxrconclusion_chestxray", 
                      "homefuel", "hivstatus", "tb")
numerical_vars <- c("Age", "oxygencons")

# Set theme with white text
my_theme <- theme_minimal() +
  theme(
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white")
  )

# EDA for categorical variables
for (var in categorical_vars) {
  p <- madaproject %>%
    count(!!sym(var)) %>%
    ggplot(aes(x = fct_reorder(!!sym(var), n), y = n)) +
    geom_bar(stat = "identity", fill = "skyblue") +
    labs(title = paste("Distribution of", var), x = var, y = "Count") +
    my_theme +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  # Save the plot
  figure_file <- here::here("results", "figures", paste0(var, "_barplot.png"))
  ggsave(filename = figure_file, plot = p)
}

# EDA for numerical variables
for (var in numerical_vars) {
  p <- madaproject %>%
    ggplot(aes(x = !!sym(var))) +
    geom_histogram(fill = "skyblue", color = "grey", bins = 20) +
    labs(title = paste("Histogram of", var), x = var, y = "Frequency") +
    my_theme

}

```

# Notes

If additional steps need to be done please let me and fix the issues

### 2024-03-27 code/processing-code/processingfile-v1.qmd
project

---
title: "Predicting of the pulmunary TB cases using clinical data"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

# Processing script

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(readr)
library(tidymodels)
```

# Data loading

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","TB data.xlsx")

#load data
# Loading the needed datasets
madaproject <- read_excel(data_location)

```

# Check data

First we can look at the codebook

```{r}
codebook <- readxl::read_excel(data_location, sheet ="codebook")
print(codebook)
```

Several ways of looking at the data

```{r}
dplyr::glimpse(madaproject)
summary(madaproject)
head(madaproject)
skimr::skim(madaproject)
```

#Cleaning
By inspecting the data as done above, we find some of the data does a missing value

Now I will at the data types

```{r}
#Now I will look at the data types of the variables
str(madaproject)
colnames(madaproject)
```
Now I see that most of the variables are characters yet most of are supposed to be categorical and i see home fuel being numeric yet its sussposed to be categorical

```{r}

# converting character variables as factors 
madaproject <- mutate_at(madaproject, 
                         vars(Gender, Asthma, Smoke, Alcohol, FeverNO, Weightloss, 
                              CoughNO, `Sputumproduction`, `BloodinSputum`, 
                              cxrconclusion_chestxray, homefuel, HIVrslt, Spatum_Culture), 
                         as.factor)

```
Then I will look and see if the structure has changed

```{r}
# now I will look at the strusture
str(madaproject)

```
since now the variables have the right data types I will look again to ensure i have no missing values

```{r}
#generally I will look at missing generally
naniar::gg_miss_var(madaproject)
```
No I see I have some missing data in the outcome variable , now i will list the nas

```{r}
# Check for missing values in each column
na_counts <- colSums(is.na(madaproject))

# Display columns with missing values
na_columns <- names(na_counts[na_counts > 0])
na_columns
```
I see I have one , I will remove it from column BloodinSputum since its one it might not cause an impact
```{r}
madaproject <- na.omit(madaproject)
```

Also I see Home Fuel has 0 and 1 so I will rename the levers to yes = 1 and No = 0

```{r}
# Rename levels
madaproject$homefuel <- ifelse(madaproject$homefuel == "1", "Yes", "No")
```

Testing to comfirm that there is no missing values.
```{r}
#generally I will look at missing generally
naniar::gg_miss_var(madaproject)
```

I will now rename some variables to make them more understable
```{r}
# Renaming the some of the variables 
madaproject <- madaproject %>% 
              rename(tb = Spatum_Culture,
                     feverstatus = FeverNO,
                     coughstatus = CoughNO,
                     hivstatus = HIVrslt)

```

Now I will check my variable
```{r}
colnames(madaproject)
```
Good  the variables changed

I feel the data is clean I will go ahead and save it has my final dataset

# Save data 
Finally, I will save the clean data as RDS file.

```{r}
# Save finaldata as RDS
save_data_location <- here::here("data","processed-data","madaproject.rds")
saveRDS(madaproject, file = save_data_location)
```

# Notes

If there is any issue with this file and the data created please let me so that I fix the issues.

### 2024-03-27 products/manuscript/Manuscript.qmd
project

---
title: "Prediction of Tuberculosis from Clinical Data"
subtitle: ""
author: Patrick Kaggwa
date: "2024-02-24"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/mada.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

{{< pagebreak >}}

# Introduction

Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@Xin2019]. This delay impacts individual prognosis and community transmission. TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@Bulled2020],. In 2020, TB ranked among the top 10 causes of death globally, with over 10 million new cases and 1.3 million deaths reported.

Africa bears a considerable burden, with an estimated quarter of new cases globally, resulting in approximately 417,000 deaths annually[@WHO2023]. TB is a leading cause of death among HIV-infected patients in sub-Saharan Africa, compounded by the low sensitivity of the commonly used sputum smear microscopy, particularly in detecting TB among people living with HIV.

## General Background Information

In Uganda, TB remains a significant public health concern, with an incidence of 330 cases per 100,000 people annually, including 136 new smear-positive cases per 100,000. Diagnostic delays persist due to health services often waiting for systematic symptoms before examining sputum smears, leading to missed opportunities for timely diagnosis.

The GeneXpert diagnostic tool offers superior sensitivity compared to sputum smear microscopy but is underutilized in Uganda due to its expense, resulting in over 41,000 undiagnosed cases annually. Even with existing diagnostic methods like blood tests or sputum tests, analysis times are prolonged, allowing culture-positive TB cases to go undetected.

## Description of data and data source

A thorough review of patients' medical records was conducted to gather data on various risk factors associated with tuberculosis (TB). This involved examining all patient forms and their radiography reports, as well as analyzing laboratory test results obtained from the clinic's records. The data collected included demographic variables such as gender and age, as well as clinical indicators such as oxygen consumption, asthma status, smoking habits, alcohol consumption, fever status, weight loss, cough status, sputum production, and presence of blood in sputum. Additionally, information on environmental factors such as the type of home fuel used was documented. Furthermore, HIV status and tuberculosis status were also recorded. This comprehensive approach to data collection ensured a thorough understanding of the factors influencing TB infection and progression within the patient population under study.

## Questions/Hypotheses to be addressed

How can TB prediction be modeled using early clinical patient data?

What is the performance of the TB prediction models? {{< pagebreak >}}

# Methods

## Study Design

The study was a retrospective study. This study was based on the secondary data, which was classified as clinical examination, patients' history like alcohol use, smoking and diagnostics.

## Study Population and Setting

The study population was the patients' medical records between 2011 to 2018 who visited the hospital under an International Development Research Centre (IDRC) project. The project is currently running in China Friendship Hospital, Naguru.

## Inclusion and Exclusion Criteria

The inclusion criteria was all patients who had cough for more than two weeks and the exclusion criteria was all adults living outside Kampala.

## Data aquisition

The data was acquired from the TB clinic at Hospital patient records form patient forms and medical examination reports from the data records office. The TB dataset consisted of 2296 instances with 15 attributes.

## Data import and cleaning

Our analysis began with thorough data preparation. Following a systematic approach, we ensured data integrity and reliability by sourcing and processing the dataset from an Excel file into R. To understand the data's structure, we reviewed the codebook and conducted exploratory analyses, including summaries and visualizations. We then cleaned the data to enhance its quality for analysis. Character variables were converted to factors, and missing values were identified using visualizations and summaries and omitted. Finally, the data was strategically split into training and testing sets in an 70/30 ratio to prepare for model development and evaluation. These steps guaranteed a dataset ready for further analysis.

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

Data analysis was conducted using R Studio version RStudio 2023.09.1. Descriptive statistics (frequencies, mean, standard deviation (SD), median, interquartile range (IQR), and proportions) were used to present the baseline characteristics of participants. we used histograms, bar graphs, and box plots, to show how data was distributed.We did cross tabulation between the covariates and the outcome.A dataset of 2294 patients was analyzed. Amongst these patients, the gender distribution was 47.0% female and 53.0% male with mean age of 31 years.

```{r}
table1 <- readRDS(here::here("results", "tables", "table1.rds"))
table_gt <- gtsummary::as_gt(table1)
table_gt |> gt::as_raw_html()
```

Figure1. Age distribution by TB

```{r}
#Figure1. Age distribution by TB
knitr::include_graphics(here("results","figures", "age_boxplot.png"))
```

figure2. TB and Age

## Basic statistical analysis

Model 1 which does not have cross validation has a very low ROC AUC value (12.21%), indicating poor discrimination between positive and negative instances. This suggests that Model 1's performance in distinguishing between the two classes is not much better than random guessing. Model 2 which uses cross validation, on the other hand, exhibits a significantly higher ROC AUC value (86.78%), suggesting much better performance in distinguishing between the positive and negative instances. A value of 86.78% indicates strong discriminatory power, with higher true positive rates and lower false positive rates across various thresholds. talk about the model number of folds

#Figure1. Age distribution by TB knitr::include_graphics(here("results","figures", "Roccurve.png"))

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

{{< pagebreak >}}

# References

### 2024-03-27 code/analysis-code/statistical-analysis.R

@@ -5,49 +5,82 @@
 #and saves the results to the results folder
 
 #load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
+library(here) #for data loading/saving#load needed packages. make sure they are installed.
+library(tidyverse) #for data processing/cleaning; includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
+library(skimr) #for nice visualization of data 
+library(here) #to set paths
+library(renv) #for package management
+library(knitr) #for nice tables
+library(kableExtra) #for nice tables
+library(gt) #for nice tables
+library(gtsummary) #for summary tables
+
 
 #path to data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
+data_location <- here::here("data","processed-data","madaproject.rds")
 
 #load data. 
-mydata <- readRDS(data_location)
+madaproject <- readRDS(data_location)
 
-
-######################################
-#Data fitting/statistical analysis
-######################################
 
 ############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
+## ---- fit-data-logistic --------
 
-lmfit1 <- lm(Height ~ Weight, mydata)  
+# second model has all variables as predictors
+log_mod <- logistic_reg() %>% set_engine("glm")
+logfit <- log_mod %>% fit(tb ~ ., data = madaproject)
 
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
+# Load required libraries
+library(tidymodels)
+library(pROC)
 
-#look at fit results
-print(lmtable1)
+# Define the logistic regression model with glm engine
+log_mod <- logistic_reg() %>% 
+  set_engine("glm")
 
-# save fit results table  
-table_file1 = here("results", "tables", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+# Fit the model
+logfit <- log_mod %>% 
+  fit(tb ~ ., data = madaproject)
 
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
+# Predict probabilities
[…]

### 2024-04-07 code/analysis-code/ML-analysis.qmd
part 4

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

## Introduction

This script uses tuberculosis(tb) as the outcome of interest and fits the following models to the analysis data:

* Null
* LASSO
* Random Forest
* Support Vector Machine

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Each model will follow this process:
1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

```{r seed=123}
defaultW <- getOption("warn") 
options(warn = -1) 
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(yardstick)
library(probably) #install.packages('probably')
tidymodels_prefer()
library(kernlab) 
options(warn = defaultW)

```

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","data2.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)
head(madaproject)
colnames(madaproject)
```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
# evaluate class imbalance 
madaproject %>% count(tb)
```
I see that i have more people without tb

Let me look at the levels 
```{r}
levels(madaproject$tb)
```
I see that NO is predicted first but I want yes

```{r}
# If "YES" is not already the first level, use relevel() to interchange the levels
madaproject$tb <- relevel(madaproject$tb, ref = "YES")

# Check the levels again to confirm the change
levels(madaproject$tb)
```
Now I see that YES is the first level

Now I will split the data into training and testing 
```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)

# Data splits , train, test
data_split <- initial_split(madaproject, prop = 0.7)
train <- training(data_split)
test <- testing(data_split)
ntrain = nrow(train)
ntest = nrow(test)
```

FITTING OF ALL MODELS 
Fitting the null model
```{r}
#Fit a null model to the outcome tb
logistic_model <- logistic_reg() %>% set_engine("glm")
null_model <- logistic_model %>% fit(tb ~ 1, data = train)

#Compute acurrancy  for the null model
metrics_null <- null_model %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, .pred_class)
#Print accurancy for the null model
print(metrics_null)

```
LASSO MODEL TUNING

```{r}
set.seed(rngseed)
# Make sure you set reference level (the outcome you are NOT interested in)
levels(madaproject$tb)

data_cv10 <- vfold_cv(train, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(tb ~ ., data = train) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_recipe(logistic_rec) %>%
    add_model(logistic_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
lasso_grid <- grid_regular(
  penalty(range = c(-3, 1)), 
  levels = 30)

tune_outputlasso <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE),
  grid = lasso_grid # lasso penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_outputlasso) + theme_classic()

# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_outputlasso, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the lowest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_lasso <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
fit(data = train)

```

```{r}
set.seed(rngseed)
# Predict probabilities
lassomodel_probs <- final_fit_lasso %>%
  predict(new_data = train, type = "prob")  
lassomodel_with_probs <- bind_cols(train, lassomodel_probs)

#Calculate ROC curve
roc_curve_lasso <- roc_curve(data = lassomodel_with_probs, truth = tb, .pred_YES)

Lasso_ROCCURVE <- autoplot(roc_curve_lasso)
Lasso_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "Lasso_ROCCURVE.png")
ggsave(filename = figure_file, plot=Lasso_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(final_fit_lasso, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matlasso <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
lassoconf_table<- as.data.frame(conf_matlasso$table)

# Create a flextable from the data frame
#lassoconf_table <- flextable(conf_dflasso)

# Print the table
print(lassoconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "lassoconf_table.rds")
saveRDS(lassoconf_table, file = summarytable_file)
```

Now I will do Random Forest
```{r}
# Random Forest Model Spec
rf_spec_tune <- rand_forest() %>%
    set_mode("classification") %>%
    set_engine("ranger") # or any other suitable engine

# Recipe
rf_rec <- recipe(tb ~ ., data = train)

# Workflow (Recipe + Model)
rf_wf <- workflow() %>% 
    add_recipe(rf_rec) %>%
    add_model(rf_spec_tune)

# Tune Model (trying a variety of values of mtry and trees)
rf_grid <- grid_regular(
  mtry(range = c(1,7)), # adjusting based on number of predictors
  trees(range = c(1, 21)), # adjust the range as needed
  levels = 5 # adjust the number of levels as needed
)

tune_outputrf <- tune_grid( 
  rf_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc),
  control = control_resamples(save_pred = TRUE),
  grid = rf_grid, # parameter grid defined above
)
print(tune_outputrf)
# Visualize Model Evaluation Metrics from Tuning
#autoplot(tune_outputrf)

# Fit Final Model
final_fit_rf <- finalize_workflow(rf_wf, select_best(tune_outputrf, "roc_auc")) %>% 
    fit(data = train)

```

```{r}

set.seed(rngseed)
# Predict probabilities
rfmodel_probs <- final_fit_rf %>%
  predict(new_data = train, type = "prob")  
rfmodel_with_probs <- bind_cols(train, rfmodel_probs)

#Calculate ROC curve
roc_curve_rf <- roc_curve(data = rfmodel_with_probs, truth = tb, .pred_YES)

rf_ROCCURVE <- autoplot(roc_curve_rf)
rf_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "rf_ROCCURVE.png")
ggsave(filename = figure_file, plot=rf_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(final_fit_rf, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matrf <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
rfconf_table <- as.data.frame(conf_matrf$table)

# Create a flextable from the data frame
#rfconf_table <- flextable(conf_dfrf)

# Print the table
print(rfconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "rfconf_table.rds")
saveRDS(rfconf_table, file = summarytable_file)
```

Using Support Vector Machine 
```{r}
# Support Vector Machine Model Spec
svm_spec_tune <- svm_poly() %>%
    set_mode("classification") %>%
    set_engine("kernlab") # or any other suitable engine for SVM

# Recipe
svm_rec <- recipe(tb ~ ., data = train)

# Workflow (Recipe + Model)
svm_wf <- workflow() %>% 
    add_recipe(svm_rec) %>%
    add_model(svm_spec_tune)

# Tune Model (trying a variety of values of C and degree for polynomial kernel)
svm_grid <- grid_regular(
  cost(range = c(0.1, 10)), # adjust the range as needed
  degree(range = c(1, 5)), # adjust the range as needed
  levels = 5 # adjust the number of levels as needed
)

tune_outputsvm <- tune_grid( 
  svm_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc),
  control = control_resamples(save_pred = TRUE),
  grid = svm_grid # parameter grid defined above
)

# Fit Final Model
svmfinal_fit <- finalize_workflow(svm_wf, select_best(tune_outputsvm , "roc_auc")) %>% 
    fit(data = train)

```

```{r}
set.seed(rngseed)
# Predict probabilities
svmmodel_probs <- svmfinal_fit %>%
  predict(new_data = train, type = "prob")  
svmmodel_with_probs <- bind_cols(train, svmmodel_probs)

#Calculate ROC curve
roc_curve_svm <- roc_curve(data = rfmodel_with_probs, truth = tb, .pred_YES)

svm_ROCCURVE <- autoplot(roc_curve_svm )
svm_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "svm_ROCCURVE.png")
ggsave(filename = figure_file, plot=svm_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(svmfinal_fit, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matsvm <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
svmconf_table <- as.data.frame(conf_matsvm$table)

# Create a flextable from the data frame
#svmconf_table <- flextable(conf_dfsvm)

# Print the table
print(svmconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "svmconf_table.rds")
saveRDS(svmconf_table, file = summarytable_file)

```
Organising results of accuracy and RMSE

```{r}
# Null model
# Compute the accuracy and AUC for null model
null_acc <- null_model %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

null_auc <-  null_model %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#LASSO model
# Compute the accuracy and AUC for Lasso
lasso_acc <- final_fit_lasso %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

lasso_auc <-  final_fit_lasso %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
rf_acc <- final_fit_rf %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

rf_auc <-  final_fit_rf %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Support Vector Machine
# Compute the accuracy and AUC for SVM
svm_acc <- svmfinal_fit %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

svm_auc <-  svmfinal_fit %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)

```

```{r}
# Print the results
print(null_acc)
print(lasso_acc)
print(rf_acc)
print(svm_acc)
print(null_auc)
print(lasso_auc)
print(rf_auc)
print(svm_auc)
```
```{r}
# Create a data frame with model names, accuracy, and AUC
model_perfomance_train <- data.frame(
  Model = c("Null Model", "Lasso Model", "Random Forest Model", "Support Vector model"),  # Replace with your model names
  Accuracy = c(0.5781931	, 0.7831776	, 0.8255452	, 0.7838006	),  # Replace with your accuracy values
  AUC = c(0.5	, 0.8707883	, 0.8973205		, 0.8618063		)  # Replace with your AUC values
)

# Print the results
print(model_perfomance_train)

# Save the table as a file

summarytable_file <- here("results","tables", "model_perfomance_train.rds")
saveRDS(model_perfomance_train, file = summarytable_file)

```

MODEL EVALUTIONS
I will fit the models again using test data and look at the results

```{r}

# Null model
# Compute the accuracy and AUC for null model
null_acc <- null_model %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

null_auc <-  null_model %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#LASSO model
# Compute the accuracy and AUC for model 2
lasso_acc <- final_fit_lasso %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

lasso_auc <-  final_fit_lasso %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
rf_acc <- final_fit_rf %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

rf_auc <-  final_fit_rf %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
svm_acc <- svmfinal_fit %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

svm_auc <-  svmfinal_fit %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)

```

```{r}
# Print the results
print(null_acc)
print(lasso_acc)
print(rf_acc)
print(svm_acc)
print(null_auc)
print(lasso_auc)
print(rf_auc)
print(svm_auc)
```

```{r}
# Create a data frame with model names, accuracy, and AUC Based on test data
model_perfomance_test <- data.frame(
  Model = c("Null Model", "Lasso Model", "Random Forest Model", "Support Vector model"),  # Replace with your model names
  Accuracy = c(0.6023222	, 0.8026125	, 0.8156749		, 0.796807),  # Replace with your accuracy values
  AUC = c(0.5	, 0.8894864		, 0.8868349			, 0.882473			)  # Replace with your AUC values
)

# Print the results
print(model_perfomance_test)

# Save the table as a file

summarytable_file = here("results","tables", "model_perfomance_test.rds")
saveRDS(model_perfomance_test, file = summarytable_file)

```

### 2024-04-07 code/analysis-code/chi-square-testin.qmd
part 4

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

```{r}
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(rsample)
library(vcd)# Load the vcd package for Cramer's V calculation
```

The purpose of this check with variables are related to the outcome and will run chi squares to check the p values

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)

```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
colnames(madaproject)
```

```{r}

# Selecting the relevant columns
relevant_columns <- c("Gender", "Age", "oxygencons", "Asthma", "Smoke", 
                      "Alcohol", "feverstatus", "Weightloss", "coughstatus", 
                      "Sputumproduction", "BloodinSputum", "cxrconclusion_chestxray", 
                      "homefuel", "hivstatus", "tb")

# Subset the data with relevant columns
relevant_data <- madaproject[, relevant_columns]

# Convert 'tb' to a factor if it's not already
relevant_data$tb <- factor(relevant_data$tb)

# Perform chi-square test for each variable against 'tb'
chi_square_results <- lapply(names(relevant_data)[-which(names(relevant_data) == "tb")], 
                              function(var) {
                                  chisq.test(relevant_data[[var]], relevant_data$tb)
                              })

# Print the results
names(chi_square_results) <- names(relevant_data)[-which(names(relevant_data) == "tb")]
chi_square_results

```

Now I will create a dataset that variables whose p value are less than 0.05

```{r}
# Selecting the variables with p- values less than 0.05
data2 <- madaproject %>% select( "oxygencons", "Asthma", 
                       "feverstatus", "Weightloss", 
                       "BloodinSputum", "cxrconclusion_chestxray", 
                      "homefuel", "hivstatus", "tb")

colnames(data2)
```

Running again chisquare test among each variable
```{r}

# contingency tables and perform chi-square tests for each pair of variables

# Create an empty list to store results
results <- list()

# Iterate through each pair of variables
for (i in 1:(length(madaproject) - 1)) {
  for (j in (i + 1):length(data2)) {
    # Create contingency table
    contingency_table <- table(madaproject[[i]], data2[[j]])

    # Perform chi-square test
    chi_square_result <- chisq.test(contingency_table)

    # Calculate Cramer's V
    cramer_v <- assocstats(contingency_table)$cramer

    # Store results in the list
    results[[paste(names(data2)[i], names(madaproject)[j], sep="_vs_")]] <- list(chi_square_result = chi_square_result, cramer_v = cramer_v)
  }
}

# Print results
for (result_name in names(results)) {
  cat("Results for", result_name, ":\n")
  print(results[[result_name]])
  cat("\n")
}

```

```{r}
# Load necessary libraries
library(vcd)
library(flextable)

# Create an empty data frame to store results
results_df <- data.frame(
  Variable1 = character(),
  Variable2 = character(),
  ChiSquarePValue = numeric(),
  CramersV = numeric(),
  stringsAsFactors = FALSE
)

# Iterate through each pair of variables
for (i in 1:(length(madaproject) - 1)) {
  for (j in (i + 1):length(madaproject)) {
    # Create contingency table
    contingency_table <- table(madaproject[[i]], madaproject[[j]])

    # Perform chi-square test
    chi_square_result <- chisq.test(contingency_table)

    # Calculate Cramer's V
    cramer_v <- assocstats(contingency_table)$cramer

    # Store results in the data frame, rounding values to three decimal places
    results_df <- rbind(results_df, 
                        data.frame(
                          Variable1 = names(madaproject)[i],
                          Variable2 = names(madaproject)[j],
                          ChiSquarePValue = sprintf("%.3f", chi_square_result$p.value),
                          CramersV = sprintf("%.3f", cramer_v)
                        )
    )
  }
}

# Create a flextable object
ft <- flextable(results_df)

# Print the flextable
ft

```

Now I have only 9 predictors
I will now save the dataset

```{r}
# Save finaldata as RDS
save_data_location <- here::here("data","processed-data","data2.rds")
saveRDS(data2, file = save_data_location)
```

### 2024-04-07 code/analysis-code/ordinary_logistic-analysis.qmd
part 4

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

## Introduction

This script uses tuberculosis(tb) as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* Random Forest
* LASSO

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

```{r seed=123}
defaultW <- getOption("warn") 
options(warn = -1) 
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(yardstick)
options(warn = defaultW)

```

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)
colnames(madaproject)
```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)
data_split <- initial_split(madaproject, prop = 0.7)
train <- training(data_split)
test <- testing(data_split)
ntrain = nrow(train)
ntest = nrow(test)
```

## ---- fit-data-logistic --------
# fit the logistic models with tb as outcome 
# second model has all variables as predictors
```{r}
log_mod <- logistic_reg() %>% set_engine("glm")
logmod2 <- log_mod %>% fit(tb ~ ., data = train)

```

```{r}
# Compute the accuracy and AUC for model 2
logmod2_acc <- logmod2 %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))
```

```{r}
levels(train$tb)
```

```{r}

# Predict probabilities
logmod2_probs <- logmod2 %>%
  predict(new_data = train, type = "prob")  
```

```{r}
logmod2_auc <-  logmod2 %>%
  predict(new_data = train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
# Print the results
print(logmod2_acc)
print(logmod2_auc)
```

```{r}
# Bind predicted probabilities to the training data
logmod2_with_probs <- bind_cols(train, logmod2_probs)

#Save training data and predicted values as RDS
save_data_location <- here::here("data","processed-data","logmod2_with_probs.rds")
saveRDS(logmod2_with_probs, file = save_data_location)
```

```{r}
 #Calculate ROC curve
roc_curve_data <- yardstick::roc_curve(data = logmod2_with_probs, truth = tb, 
                                       .pred_YES)
#save the ROC curve as RDS

# Print ROC curve data
print(roc_curve_data)
```
```{r}
# Plot ROC curve
Roccurve <- ggplot(roc_curve_data, aes(x = sensitivity , y = 1 - specificity)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "blue") +
  labs(x = "False Positive Rate (TB CASES)", 
       y = "True Positive Rate (TB CASES)", 
       title = "ROC Curve") +
  theme_minimal()
Roccurve

figure_file = here("results","figures", "Roccurve.png")
ggsave(filename = figure_file, plot=Roccurve) 
```

#Model improvement

```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)

#Define cross-validation with k = 10 folds
folds <- vfold_cv(train, v = 10)

#setting model specification 
model_specf <- logistic_reg() %>% set_engine("glm")

# Workflow for all predictors
wfall4 <- workflow() %>%
  add_model(model_specf)%>%
  add_formula(tb ~ .,)

#Fitting the models with resampling 
fit2 <- fit_resamples(wfall4, resamples = folds)

#Getting metrics for both models 
collect_metrics(fit2)
```

`

### 2024-04-07 code/eda-code/eda.qmd
part 4

@@
 ```
+Creating a table of characteristics
 
 ```{r}
+library(dplyr)
+library(gtsummary)
 
-# Create gtsummary table 
-table1 <- 
-  madaproject %>%
-  tbl_summary(
-    include = all_of(c("Gender", "Asthma", "Smoke", "Alcohol", "feverstatus", 
+# Convert non-numeric variables to character
+madaproject <- madaproject %>%
+  mutate(across(where(is.factor), as.character))
+
+# Define variables of interest (excluding "Age" and "oxygencons")
+variables <- c("Gender", "Asthma", "Smoke", "Alcohol", "feverstatus", 
                "Weightloss", "coughstatus", "Sputumproduction", 
                "BloodinSputum", "cxrconclusion_chestxray", 
-                       "homefuel", "hivstatus", "tb")),
-    missing = "no",
-    label = list(
-      Asthma ~ "Asthma",
-      Smoke ~ "Smoking Status",
-      Alcohol ~ "Alcohol",
-      feverstatus ~ "Fever Status",
-      Weightloss ~ "Weight Loss",
-      coughstatus ~ "Cough Status",
-      `Sputumproduction` ~ "Sputum Production",
-      `BloodinSputum` ~ "Blood in Sputum",
-      cxrconclusion_chestxray ~ "CXR Conclusion",
-      homefuel ~ "Home Fuel",
-      hivstatus ~ "HIV Status",
-      tb ~ "TB"
+               "homefuel", "hivstatus")
+
+# Create a summary table with specified variable types
+summary_table_characteristics <- madaproject %>%
+  select(all_of(variables), tb) %>%
+  tbl_summary(
+    by = tb,
+    type = everything() ~ "categorical"
   )
-  ) %>%
-  add_n()
 
-# Print the table
-table1
+# Print the summary table
+summary_table_characteristics
 
 # Save table
-saveRDS(table1, file = here::here("results", "tables", "table1.rds"))
+saveRDS(table1, file = here::here("results", "tables", "summary_table_characteristics .rds"))
 
 ```
+
 This is just my code to use a for loop 

### 2024-04-07 products/manuscript/Manuscript.qmd
part 4

@@
 library(knitr)
+library(magrittr)
+library(flextable)
+library(htmltools)
 ```
 
-# Summary/Abstract
+# Abstract
 
-{{< pagebreak >}}
+#Introduction Although early detection and treatment of Tuberculosis cases are the hallmark of successful TB control, diagnostic delay is still long and common in Uganda.
+Therefore, the study aimed atdeveloping a tool that uses machine-learning techniques to detect and diagnose TB more accurately in a shorter time.
+#Methods This was a retrospective study that used secondary data collected between 2011 to 2018 collected from health under Infectious diseases research collaboration an (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
+The study was conducted on data from 2296 tested patients of whom 1345 had no TB and 951 had TB.
+Different machine learning metrics(Accuracy, RMSE and AUC) were used evaluate the performance of the models.
+#Results {{< pagebreak >}}
 
@@
 Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@Xin2019].
-This delay impacts individual prognosis and community transmission.
+This delay impacts individual prognosis and community transmission[@f_verbeke_4_2012].
 TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@Bulled2020],.
@@
 
-What is the performance of the TB prediction models? {{< pagebreak >}}
+What is the performance of the TB prediction models?
+
+{{< pagebreak >}}
 
@@
 ```{r}
-table1 <- readRDS(here::here("results", "tables", "table1.rds"))
-table_gt <- gtsummary::as_gt(table1)
-table_gt |> gt::as_raw_html()
+#| label: tbl-characteristics
+#| tbl-cap: " Participants' Characteristics"
+#| echo: FALSE
+ch<- readRDS(here("results", "tables", "characteristics.rds"))
+
+ch %>%
+  kable()
 ```
 
-Figure1. Age distribution by TB
+##Determining requirements for Tuberculosis prediction that is appropriate data and modeling techniques
+
+In our study, we utilized 2296 instances and examined 14 variables associated with TB.
+Subsequently, nine variables were selected for prediction after conducting chi-square tests.
+Non-statistically significant variables associated with TB were eliminated.
+Following this, chi-square tests were applied to the nine chosen variables to assess their association with a threshold of 0.05.
+However, these tests did not reveal a significant association among the variables themselves.
+Therefore, they were included in the prediction model.
+The variables retained for prediction were: asthma, alcohol use, fever, chest radiography, home fuel use, weight loss, presence of blood in sputum, oxygen saturation, and HIV status.
 
 ```{r}
-#Figure1. Age distribution by TB
-knitr::include_graphics(here("results","figures", "age_boxplot.png"))
+#| label: tbl-schematic2b
+#| tbl-cap: "Model_perfomance."
+#| echo: FALSE
+mpf <- readRDS(here("results", "tables", "model_perfomance_train.rds"))
+
+mpf %>%
+  kable()
 ```
 
-figure2. TB and Age
+This study examines the efficacy of four distinct machine learning models in predicting the presence of tuberculosis (TB) using input data.
+Each model, including LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, is rigorously evaluated based on key performance metrics.
+The LASSO model applies the Least Absolute Shrinkage and Selection Operator regularization technique, while the Random Forest model employs ensemble learning.
+SVM seeks to optimize hyperplanes for class separation, and the Null model serves as a simplistic baseline.
+Performance assessment encompasses accuracy (with LASSO achieving 80.26%, Random Forest achieving 81.57%, SVM achieving 79.68%, and the Null model achieving 60.23%), Area Under the Curve (AUC) (with LASSO achieving 0.8895, Random Forest achieving 0.8868, SVM achieving 0.8825, and the Null model achieving 0.5).
+The findings underscore the potential of advanced machine learning algorithms, particularly Random Forest and LASSO, in enhancing TB diagnosis and early detection strategies.
 
-## Basic statistical analysis
+```{r}
+#| label: tbl-Random Forest
+#| tbl-cap: " Model Predictions (Random Forest)"
+#| echo: FALSE
+rf <- readRDS(here("results", "tables", "rfconf_table.rds"))
 
-Model 1 which does not have cross validation has a very low ROC AUC value (12.21%), indicating poor discrimination between positive and negative instances.
-This suggests that Model 1's performance in distinguishing between the two classes is not much better than random guessing.
-Model 2 which uses cross validation, on the other hand, exhibits a significantly higher ROC AUC value (86.78%), suggesting much better performance in distinguishing between the positive and negative instances.
-A value of 86.78% indicates strong discriminatory power, with higher true positive rates and lower false positive rates across various thresholds.
-talk about the model number of folds
+rf %>%
+  kable()
+```
 
-#Figure1. Age distribution by TB knitr::include_graphics(here("results","figures", "Roccurve.png"))
+The table presents the classification results of a Random Forest model for predicting tuberculosis (TB) status in individuals.
+Among individuals with TB (TB=YES), the model correctly identified 619 true positives and incorrectly predicted 219 cases as TB false negatives.
+For individuals without TB, the model accurately classified 709 cases as TB true negatives but misclassified 58 cases as TB false positives.
 
-{{< pagebreak >}}
+```{r}
+#| label: fig-rf_ROCCURVE
+#| fig-cap: "rf_ROCCURVE"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures", "rf_ROCCURVE.png"))
+```
+
+```{r}
+#| label: SVM
+#| tbl-cap: "  Model Predictions (SVM)"
+#| echo: FALSE
+svm<- readRDS(here("results", "tables", "svmconf_table.rds"))
+svm %>%
+  kable()
+```
+
+The table presents the classification results of a Support Vector Machine model for predicting tuberculosis (TB) status in individuals.
+Among individuals with TB, the model correctly identified 621 true positives and incorrectly predicted 291 cases as TB false negatives.
+For individuals without TB, the model accurately classified 637 cases as TB true negatives but misclassified 56 cases as TB positive (false positives).
+
+```{r}
+#| label: fig-svm_ROCCURVE
+#| fig-cap: "svm_ROCCURVE"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures", "svm_ROCCURVE.png"))
+```
+
+```{r}
+#| label: LASSO
+#| tbl-cap: "  Model Predictions (LASSO)"
+#| echo: FALSE
+ls<- readRDS(here("results", "tables", "lassoconf_table.rds"))
+ls %>%
+  kable()
+```
+
+The table presents the classification results of a LASSSO model for predicting tuberculosis (TB) status in individuals.
+Among individuals with TB, the model correctly identified 654 true positives and incorrectly predicted 291 cases as TB false negatives.
+For individuals without TB, the model accurately classified 603 cases as TB true negatives but misclassified 23 cases as TB false positives.
+
+```{r}
+#| label: fig-Lasso_ROCCURVE
+#| fig-cap: "Lasso_ROCCURVE"
+#| echo: FALSE 
+knitr::include_graphics(here("results","figures", "Lasso_ROCCURVE.png"))
+```
+```{r}
+#Figure1. Random Forest ROC Curve
+
+```
+
+#Models Evaluation {{< pagebreak >}}
 
 # Discussion
+
+The aim of the study was to apply machine-learning techniques in developing models for TB screening and detection in a shorter time based on patient data records to contribute to disease notification.
+The application of machine learning approaches have therefore been found to be successful in prediction [@Holzinger].
 

### 2024-04-08 code/analysis-code/ML-analysis.qmd
Part4_1

@@
 ```
+
+The evaluation of machine learning models' performance on test data is pivotal for gauging their efficacy and generalization capabilities.
+In this study, conducted to assess predictive accuracy, the models were measured using metrics such as accuracy and Area Under the Curve (AUC).
+The Lasso, Random Forest, and Support Vector models were compared against a Null Model, with their respective performance metrics outlined.
+The results revealed a notable performance gap between the advanced models and the Null Model.
+Specifically, the Random Forest model exhibited superior performance, achieving an accuracy of 81.57% and an AUC of 0.887.
+In contrast, the Lasso and Support Vector models attained accuracies of 80.26% and 79.68%, with corresponding AUC values of 0.889 and 0.882, respectively.

### 2024-04-08 code/eda-code/eda.qmd
Part4_1

@@
 # Create a summary table with specified variable types
-summary_table_characteristics <- madaproject %>%
+characteristics <- madaproject %>%
   select(all_of(variables), tb) %>%
@@
   )
-
+characteristics <- data.frame(characteristics)
 # Print the summary table
-summary_table_characteristics
+characteristics
 
 # Save table
-saveRDS(table1, file = here::here("results", "tables", "summary_table_characteristics .rds"))
+
+summarytable_file <- here("results", "tables", "characteristics.rds")
+saveRDS(characteristics, file = summarytable_file)
 

### 2024-04-08 products/manuscript/Manuscript.qmd
Part4_1

@@
 ---
-title: "Prediction of Tuberculosis from Clinical Data"
+title: "CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING"
 subtitle: ""
@@
 
-#Introduction Although early detection and treatment of Tuberculosis cases are the hallmark of successful TB control, diagnostic delay is still long and common in Uganda.
+## Introduction
+
+Although early detection and treatment of Tuberculosis cases are the hallmark of successful TB control, diagnostic delay is still long and common in Uganda.
 Therefore, the study aimed atdeveloping a tool that uses machine-learning techniques to detect and diagnose TB more accurately in a shorter time.
-#Methods This was a retrospective study that used secondary data collected between 2011 to 2018 collected from health under Infectious diseases research collaboration an (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
+
+## Methods
+
+This was a retrospective study that used secondary data collected between 2011 to 2018 collected from health under Infectious diseases research collaboration an (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
 The study was conducted on data from 2296 tested patients of whom 1345 had no TB and 951 had TB.
 Different machine learning metrics(Accuracy, RMSE and AUC) were used evaluate the performance of the models.
-#Results {{< pagebreak >}}
+
+## Results
+Evaluation of four machine learning models—LASSO, Random Forest, Support Vector Machine (SVM), and Null model—revealed notable performance differences.
+Notably, the Random Forest model achieved the highest accuracy at 81.57% and an AUC of 0.8868, outperforming the other models.
+
+## Consclusion
+
+These results highlight the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies, offering valuable insights for clinical decision-making and the development of more effective screening protocols.
+{{< pagebreak >}}
 
@@
 Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@Xin2019].
-This delay impacts individual prognosis and community transmission[@f_verbeke_4_2012].
+This delay impacts individual prognosis and community transmission[@].
 TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@Bulled2020],.
@@
 
-How can TB prediction be modeled using early clinical patient data?
+1.  How can TB prediction be modeled using early clinical patient data?
 
-What is the performance of the TB prediction models?
+2.  What is the performance of the TB prediction models?
 
@@
 
-*Explain anything related to your statistical analyses.*
-
 {{< pagebreak >}}
@@
 
-```{r}
+```{r }
 #| label: tbl-characteristics
@@
 
-##Determining requirements for Tuberculosis prediction that is appropriate data and modeling techniques
+## Determining requirements for Tuberculosis prediction that is appropriate data and modeling techniques
 
@@
 
-```{r}
-#| label: tbl-schematic2b
+```{r }
+#| label: tbl-model_perfomance_train
 #| tbl-cap: "Model_perfomance."
@@
 
-```{r}
+```{r }
 #| label: tbl-Random Forest
@@
 
-```{r}
-#| label: fig-rf_ROCCURVE
-#| fig-cap: "rf_ROCCURVE"
+```{r }
+#| label: fig-Random Forest_Roccurve
+#| fig-cap: "Random Forest_Roccurve"
 #| echo: FALSE
@@
 
-```{r}
-#| label: SVM
+```{r }
+#| label: tbl-Support Vector Machine
 #| tbl-cap: "  Model Predictions (SVM)"
@@
 
-```{r}
-#| label: fig-svm_ROCCURVE
+```{r }
+#| label: fig-supporter Vector Machine_Roccurve
 #| fig-cap: "svm_ROCCURVE"
@@
 
-```{r}
-#| label: LASSO
+```{r }
+#| label: tbl-LASSO
 #| tbl-cap: "  Model Predictions (LASSO)"
@@
 
-```{r}
+```{r }
 #| label: fig-Lasso_ROCCURVE
@@
 ```
+
 ```{r}
-#Figure1. Random Forest ROC Curve
+#| label: model_perfomance_test
+#| tbl-cap: "model_perfomance_test"
+#| echo: FALSE
+tmpf <- readRDS(here("results", "tables", "model_perfomance_test.rds"))
 
+tmpf %>%
+  kable()
 ```
 
-#Models Evaluation {{< pagebreak >}}
+# Models Evaluation
+
+The evaluation of machine learning models' performance on test data is pivotal for gauging their efficacy and generalization capabilities.
+The predictive accuracy of the models were measured using the test data.
+The Lasso, Random Forest, and Support Vector models were compared against a Null Model, with their respective performance metrics outlined.
+The results revealed a notable performance gap between the advanced models and the Null Model.
+Specifically, the Random Forest model exhibited superior performance, achieving an accuracy of 81.57% and an AUC of 0.887.
+In contrast, the Lasso and Support Vector models attained accuracies of 80.26% and 79.68%, with corresponding AUC values of 0.889 and 0.882, respectively.
+
+{{< pagebreak >}}
 
@@
 
-The aim of the study was to apply machine-learning techniques in developing models for TB screening and detection in a shorter time based on patient data records to contribute to disease notification.
-The application of machine learning approaches have therefore been found to be successful in prediction [@Holzinger].
+The study presents a comprehensive analysis of machine learning models' efficacy in predicting the presence of tuberculosis (TB) based on a dataset comprising 2296 instances and 14 variables associated with TB.
+The study selected nine variables for prediction after conducting chi-square tests to eliminate non-statistically significant variables associated with the outcome.
+The retained variables for prediction included asthma, alcohol use, fever, chest radiography, home fuel use, weight loss, presence of blood in sputum, oxygen saturation, and HIV status.
+The study evaluated four distinct machine learning models, namely LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, for their predictive performance using key metrics accuracy and Area Under the Curve (AUC).
+The LASSO model achieved an accuracy of 80.26% and an AUC of 0.8895.
+The Random Forest model outperformed others with an accuracy of 81.57% and an AUC of 0.8868.
+SVM achieved an accuracy of 79.68% and an AUC of 0.8825.
+The Null model achieved an accuracy of 60.23% and an AUC of 0.5.
 
-## Summary and Interpretation
+Furthermore, the study presented detailed classification results for each model, emphasizing their ability to correctly classify true positives and true negatives while minimizing false positives and false negatives.
+Particularly, the Random Forest model demonstrated robust performance in correctly identifying true positives and true negatives, with minimal misclassifications.
+The findings underscore the potential of advanced machine learning algorithms, particularly Random Forest and LASSO, in enhancing TB diagnosis and early detection strategies.
+The manuscript provides valuable insights into the predictive capabilities of different machine learning models which can inform clinical decision-making and aid in the development of more effective TB screening tolls and protocols.
 
-## Strengths and Limitations
+# Strengths
 
-## Conclusions
+In this study, a rigorous variable selection process was implemented, utilizing chi-square tests to identify statistically significant predictors associated with tuberculosis (TB) from a dataset comprising 2296 instances and 14 variables.
+By employing stringent selection criteria, the study ensures that only variables with meaningful associations with TB are included in subsequent analyses, thereby enhancing the model's interpretability, and reducing the risk of overfitting.
+Moreover, the evaluation of multiple machine learning models, including LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, contributes to a comprehensive comparison of different predictive approaches.
+This assessment allows for informed decision-making regarding model selection, facilitating the identification of the most suitable algorithm for TB prediction.
+
+Additionally, the study employs robust performance metrics such as accuracy and Area Under the Curve (AUC) to evaluate model performance objectively.
+These metrics provide a comprehensive assessment of each model's predictive capabilities, enabling comparisons across different algorithms.
+Furthermore, the detailed classification results presented offer insights into the models' ability to correctly classify true positives and true negatives while minimizing false positives and false negatives.
+
+# Limitations and Recommendations
+
+Despite the strengths, this study had limitations that warrant consideration.
+Firstly, the dataset's potential class imbalance, characterized by variations in the number of instances for individuals with and without TB, may impact model performance of the model.
+Additionally, while the variable selection methodology employed chi-square tests, alternative feature selection methods like recursive feature elimination could offer further insights into the predictive variables' relevance and importance.
+Future research should aim incorporate larger, more diverse datasets and conducting external validation studies to assess model performance in real-world clinical settings and use of more robust techniques such as neural networks.
+
+# Conclusions
+
+The findings underscore the potential of machine learning algorithms, particularly Random Forest and LASSO, in enhancing TB diagnosis and early detection strategies.
+Future research should further advance the development of effective TB screening protocols and contribute to improved TB diagnosis and management on a global scale.
 

### 2024-04-08 products/manuscript/Manuscript.qmd
part4_2

@@
 
-These results highlight the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies, offering valuable insights for clinical decision-making and the development of more effective screening protocols.
+The results highlight the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies, offering valuable insights for clinical decision-making and the development of more effective screening protocols.
 {{< pagebreak >}}

### 2024-04-22 products/manuscript/Manuscript.qmd
improving 1

@@
     highlight-style: github
-bibliography: ../../assets/mada.bib
+bibliography: ../../assets/MADAPROJECT.bib
 csl: ../../assets/apa.csl
@@
 
-Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@Xin2019].
+Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@sagili_strategies_2022].
 This delay impacts individual prognosis and community transmission[@].
-TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@Bulled2020],.
+TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@baykan_extrapulmonary_2022],.
 In 2020, TB ranked among the top 10 causes of death globally, with over 10 million new cases and 1.3 million deaths reported.
-
-Africa bears a considerable burden, with an estimated quarter of new cases globally, resulting in approximately 417,000 deaths annually[@WHO2023].
-TB is a leading cause of death among HIV-infected patients in sub-Saharan Africa, compounded by the low sensitivity of the commonly used sputum smear microscopy, particularly in detecting TB among people living with HIV.
+Africa bears a considerable burden, with an estimated quarter of new cases globally, resulting in approximately 417,000 deaths annually[@chakaya_who_2022].
+TB is a leading cause of death among HIV-infected patients in sub-Saharan Africa, compounded by the low sensitivity of the commonly used sputum smear microscopy, particularly in detecting TB among people living with HIV[@palattiyil_access_2022].
 
@@
 
-In Uganda, TB remains a significant public health concern, with an incidence of 330 cases per 100,000 people annually, including 136 new smear-positive cases per 100,000.
+In Uganda, TB remains a significant public health concern, with an incidence of 330 cases per 100,000 people annually, including 136 new smear-positive cases per 100,000[@kintu_unfavorable_2023].
 Diagnostic delays persist due to health services often waiting for systematic symptoms before examining sputum smears, leading to missed opportunities for timely diagnosis.
-
-The GeneXpert diagnostic tool offers superior sensitivity compared to sputum smear microscopy but is underutilized in Uganda due to its expense, resulting in over 41,000 undiagnosed cases annually.
-Even with existing diagnostic methods like blood tests or sputum tests, analysis times are prolonged, allowing culture-positive TB cases to go undetected.
+The GeneXpert diagnostic tool offers superior sensitivity compared to sputum smear microscopy but is underutilized in Uganda due to its expense[@brown_implementation_2021], resulting in over 41,000 undiagnosed cases annually.
+Even with existing diagnostic methods like blood tests or sputum tests, analysis times are prolonged, allowing culture-positive TB cases to go undetected[@dong_improved_2022].
 

### 2024-04-23 products/manuscript/Manuscript.qmd
improvement 2

@@
 
-```{r }
-#| label: tbl-characteristics
-#| tbl-cap: " Participants' Characteristics"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE }
+
 ch<- readRDS(here("results", "tables", "characteristics.rds"))
@@
 
-```{r }
-#| label: tbl-model_perfomance_train
-#| tbl-cap: "Model_perfomance."
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE}
+
 mpf <- readRDS(here("results", "tables", "model_perfomance_train.rds"))
@@
 
-```{r }
-#| label: tbl-Random Forest
-#| tbl-cap: " Model Predictions (Random Forest)"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE }
 rf <- readRDS(here("results", "tables", "rfconf_table.rds"))
-
 rf %>%
@@
 
-```{r }
-#| label: fig-Random Forest_Roccurve
-#| fig-cap: "Random Forest_Roccurve"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE }
 knitr::include_graphics(here("results","figures", "rf_ROCCURVE.png"))
@@
 
-```{r }
-#| label: tbl-Support Vector Machine
-#| tbl-cap: "  Model Predictions (SVM)"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE}
+
 svm<- readRDS(here("results", "tables", "svmconf_table.rds"))
@@
 
-```{r }
-#| label: fig-supporter Vector Machine_Roccurve
-#| fig-cap: "svm_ROCCURVE"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE}
+
 knitr::include_graphics(here("results","figures", "svm_ROCCURVE.png"))
@@
 
-```{r }
-#| label: tbl-LASSO
-#| tbl-cap: "  Model Predictions (LASSO)"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE}
+
 ls<- readRDS(here("results", "tables", "lassoconf_table.rds"))
@@
 
-```{r }
-#| label: fig-Lasso_ROCCURVE
-#| fig-cap: "Lasso_ROCCURVE"
-#| echo: FALSE 
+```{r echo=FALSE, message = FALSE}
+
 knitr::include_graphics(here("results","figures", "Lasso_ROCCURVE.png"))
@@
 
-```{r}
-#| label: model_perfomance_test
-#| tbl-cap: "model_perfomance_test"
-#| echo: FALSE
+```{r echo=FALSE, message = FALSE}
+
 tmpf <- readRDS(here("results", "tables", "model_perfomance_test.rds"))

### 2024-05-05 code/analysis-code/Cross_validation_Lasso.qmd
part61

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

```{r}
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(rsample)
library(vcd)# Load the vcd package for Cramer's V calculation
```

The purpose of this is to check if some variable are dropped by the lasso model so to get an insight of what variables

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)

```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
colnames(madaproject)
```

Now I do LASSO -Exploration

```{r}
set.seed(12345)
logistic_model <- logistic_reg() %>% set_engine("glm")

# Make sure you set reference level (the outcome you are NOT interested in)
levels(madaproject$tb)

data_cv10 <- vfold_cv(madaproject, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(tb ~ ., data = madaproject) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_model(logistic_lasso_spec_tune) %>%
    add_recipe(logistic_rec)

# Tune Model (trying a variety of values of Lambda penalty)
lasso_grid <- grid_regular(
  penalty(range = c(-3, 1)), 
  levels = 30)

tune_outputlasso <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE),
  grid = lasso_grid # lasso penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_outputlasso) + theme_classic()

# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_outputlasso, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the lowest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_lasso <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
fit(data = madaproject)
```

```{r}
# Get the best parameters
best_params <- show_best(tune_outputlasso, metric = "roc_auc") %>%
               slice(1)

# Finalize workflow with best parameters
final_lasso_wf <- finalize_workflow(log_lasso_wf, best_params)

# Fit model to training data
lasso_train_fit <- fit(final_lasso_wf, data = madaproject)

# Extract fitted model
fitted_model <- extract_fit_parsnip(lasso_train_fit)

# Extract selected variables
selected_variables <- tidy(fitted_model$fit) %>%
                      filter(!is.na(estimate) & estimate != 0) %>%
                      pull(term)

# Print selected variables
print(selected_variables)

```
I see that only one variable is dropped remaining with a total of 13 covariates.

### 2024-05-05 code/analysis-code/ML-analysis.qmd
part61

--- 2024-04-07 code/analysis-code/ML-analysis.qmd
@@
 #Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","data2.rds")
+data_location <- here::here("data","processed-data","madaproject.rds")
 #load data
@@
 
-FITTING OF ALL MODELS 
+## FITTING OF ALL MODELS
+
 Fitting the null model
@@
 log_lasso_wf <- workflow() %>% 
-    add_recipe(logistic_rec) %>%
-    add_model(logistic_lasso_spec_tune)
+    add_model(logistic_lasso_spec_tune) %>%
+    dd_recipe(logistic_rec)
 
@@
 
-# Tune Model (trying a variety of values of mtry and trees)
+# Setup grid tune for  Model
 rf_grid <- grid_regular(
@@
 print(tune_outputrf)
-# Visualize Model Evaluation Metrics from Tuning
-#autoplot(tune_outputrf)
 
@@
 rfconf_table <- as.data.frame(conf_matrf$table)
-
-# Create a flextable from the data frame
-#rfconf_table <- flextable(conf_dfrf)
 

### 2024-05-05 products/manuscript/Manuscript.qmd
HardinParker Review
part61

@@
 ## Introduction
-
 Although early detection and treatment of Tuberculosis cases are the hallmark of successful TB control, diagnostic delay is still long and common in Uganda.
-Therefore, the study aimed atdeveloping a tool that uses machine-learning techniques to detect and diagnose TB more accurately in a shorter time.
+Therefore, the study aimed at developing a tool that uses machine-learning techniques to detect and diagnose TB more accurately in a shorter time.
 
 ## Methods
-
 This was a retrospective study that used secondary data collected between 2011 to 2018 collected from health under Infectious diseases research collaboration an (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
@@
 ## Consclusion
-
 The results highlight the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies, offering valuable insights for clinical decision-making and the development of more effective screening protocols.
@@
 # Introduction
-
 Tuberculosis (TB) remains a significant global health challenge due to diagnostic limitations affecting both pediatric and adult populations, resulting in delayed diagnosis or misdiagnosis[@sagili_strategies_2022].
-This delay impacts individual prognosis and community transmission[@].
+This delay impacts individual prognosis and community transmission.
 TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@baykan_extrapulmonary_2022],.
@@
 ## Questions/Hypotheses to be addressed
-
-1.  How can TB prediction be modeled using early clinical patient data?
-
-2.  What is the performance of the TB prediction models?
+The prediction of TB can be modeled using early clinical patient data from Kampala residents who visited the IDRC between 2011 and 2018 and presented with a persistent cough lasting two weeks or more.
 
 {{< pagebreak >}}
-
 # Methods
-
 ## Study Design
-
 The study was a retrospective study. This study was based on the secondary data, which was classified as clinical examination, patients' history like alcohol use, smoking and diagnostics.
@@
 ## Study Population and Setting
-
 The study population was the patients' medical records between 2011 to 2018 who visited the hospital under an International Development Research Centre (IDRC) project.
@@
 ## Inclusion and Exclusion Criteria
-
 The inclusion criteria was all patients who had cough for more than two weeks and the exclusion criteria was all adults living outside Kampala.
@@
 ## Data aquisition
-
 The data was acquired from the TB clinic at Hospital patient records form patient forms and medical examination reports from the data records office.
@@
 ## Data import and cleaning
-
 Our analysis began with thorough data preparation.
@@
 
+## Model Development
+Three models were considered for the prediction analysis, and each were fit for both outcomes of interest (tuberculosis).
+The data were split such that 70% were used for training the models and the remaining 30% were used for testing the final chosen model.
+Cross-validation was used for all models, using a ten-fold resampling structure.
+The tidymodel framework was utilized in the analysis.
+
+## Model Definition
+
+There are several machine learning models suitable for addressing the research question, but this analysis primarily focused on three different types of models: Random Forest, LASSO, and Support Vector Machine (SVM).
+
+Random Forest (RF): RF is a versatile and robust ensemble learning method that aims to reduce variance by building multiple decision trees on random subsets of the data and then averaging their predictions.
+It can handle large datasets with high-dimensional feature spaces and is less prone to overfitting compared to individual decision trees[@simon_interpreting_2023].
+However, RF models are more complex and may be challenging to interpret compared to simpler models like decision trees.
+
+LASSO: LASSO is a regularization-based linear regression method that balances model complexity and goodness of fit by penalizing the absolute values of the regression coefficients.
+This encourages sparsity in the model, meaning it tends to select a subset of the most important features while shrinking the coefficients of less important ones towards zero[@zhao_model_2006].
+However, LASSO may struggle with highly correlated predictors and can only select one variable from a group of highly correlated variables, potentially ignoring important information.
+
+Support Vector Machine (SVM): SVM is a powerful supervised learning algorithm commonly used for classification tasks.
+It works by finding the optimal hyperplane that separates data points belonging to different classes with the maximum margin of separation.
+SVM is effective in high-dimensional spaces and is versatile in handling linear and non-linear relationships through different kernel functions.[@javatpoint_support_2021] However, SVM can be computationally intensive, especially with large datasets, and may require careful tuning of hyperparameters for optimal performance.
+
+These three models were selected to provide a diverse set of approaches for predicting TB classification based on early clinical patient data.
+Random Forest offers a robust ensemble learning approach, LASSO provides a sparse and interpretable linear model, and SVM offers flexibility in handling complex relationships between predictors.
+By comparing and contrasting the performance of these models, we aim to identify the most effective approach for TB prediction in our study population.
 {{< pagebreak >}}
 
+## Model Performance Evaluation
+
+A confusion matrix or contingency table was used to determine the classification performance of the various classification models.
+Given m classes, the confusion matrix is a table of at least size m by m [@riehl_hierarchical_2023].
+In a confusion matrix, if the instance is positive and it is classified as positive, it is counted as a true positive (TP); if it is classified as negative, it is counted as a false negative (FN).
+If the instance is negative and it is classified as negative, it is counted as a true negative (TN); if it is classified as positive, it is counted as a false positive (FP).
+These terms are useful when analyzing a classifier’s ability to determine accuracy based on the correct and incorrect classes produced by the confusion matrix.
+Accuracy was used as it signifies the proportion of the total number of TB patient predictions that are correct[@carvalho_machine_2019].
+True positive means the number of actual TB patients that are accurately classified, and true negatives refer to the number of patients that are correctly classified as not having TB.
+The ROC (Receiver Operating Characteristic) curve was also used, which is a graphical representation to measure the performance of the classifier system.
+It shows the trade-off between sensitivity and specificity for every possible cut-off for a test[@robinson_validating_2020].
+The area under the curve ranges from 0 to 1, where 1 implies an excellent test and 0 implies a useless test.
+{{< pagebreak >}}
 # Results
@@
 
-Data analysis was conducted using R Studio version RStudio 2023.09.1.
+Data analysis was conducted using R Studio version RStudio RStudio 2024.04.0.
 Descriptive statistics (frequencies, mean, standard deviation (SD), median, interquartile range (IQR), and proportions) were used to present the baseline characteristics of participants.
@@
 
-```{r echo=FALSE, message = FALSE }
-
-ch<- readRDS(here("results", "tables", "characteristics.rds"))
-
-ch %>%
-  kable()
-```
-
-## Determining requirements for Tuberculosis prediction that is appropriate data and modeling techniques
-
-In our study, we utilized 2296 instances and examined 14 variables associated with TB.
-Subsequently, nine variables were selected for prediction after conducting chi-square tests.
-Non-statistically significant variables associated with TB were eliminated.
-Following this, chi-square tests were applied to the nine chosen variables to assess their association with a threshold of 0.05.
-However, these tests did not reveal a significant association among the variables themselves.
-Therefore, they were included in the prediction model.
-The variables retained for prediction were: asthma, alcohol use, fever, chest radiography, home fuel use, weight loss, presence of blood in sputum, oxygen saturation, and HIV status.
+## Model results and accuracy assessment
 
 ```{r echo=FALSE, message = FALSE}
-
+#| label: tbl-schematic1
+#| tbl-cap: "Model Perfomance for the Training Dataset"
 mpf <- readRDS(here("results", "tables", "model_perfomance_train.rds"))
@@
 mpf %>%
-  kable()
+  knitr::kable()
 ```
@@
 ```{r echo=FALSE, message = FALSE }
+#| label: tbl-schematic2
+#| tbl-cap: "Random Forest"
 rf <- readRDS(here("results", "tables", "rfconf_table.rds"))
 rf %>%
-  kable()
+  knitr::kable()
 ```
@@
 ```{r echo=FALSE, message = FALSE }
+#| label: fig-schematic1
+#| fig-cap: "Random Forest ROC Curve"
+#| echo: false
 knitr::include_graphics(here("results","figures", "rf_ROCCURVE.png"))
@@
 ```{r echo=FALSE, message = FALSE}
-
+#| label: tbl-schematic5
+#| tbl-cap: "Support Vector Machine"
 svm<- readRDS(here("results", "tables", "svmconf_table.rds"))
 svm %>%
-  kable()
+  knitr::kable()
 ```
@@
 ```{r echo=FALSE, message = FALSE}
-
+#| label: fig-schematic2
+#| fig-cap: "Support Vector Machine ROC Curve"
+#| echo: false
 knitr::include_graphics(here("results","figures", "svm_ROCCURVE.png"))
@@
 ```{r echo=FALSE, message = FALSE}
-
+#| label: tbl-schematic3
+#| tbl-cap: "lassoconf_table"
 ls<- readRDS(here("results", "tables", "lassoconf_table.rds"))
 ls %>%
-  kable()
+  knitr::kable()
 ```
@@
 ```{r echo=FALSE, message = FALSE}
-
+#| label: fig-schematic3
+#| fig-cap: "LASSO ROC Curve"
+#| echo: false 
 knitr::include_graphics(here("results","figures", "Lasso_ROCCURVE.png"))
@@
 ```{r echo=FALSE, message = FALSE}
-
+#| label: tbl-schematic4
+#| tbl-cap: "Model Perfomance Testing Results"
 tmpf <- readRDS(here("results", "tables", "model_perfomance_test.rds"))
@@
 tmpf %>%
-  kable()
+ knitr:: kable()
 ```
@@
 In contrast, the Lasso and Support Vector models attained accuracies of 80.26% and 79.68%, with corresponding AUC values of 0.889 and 0.882, respectively.
-
 {{< pagebreak >}}

### 2024-05-05 code/eda-code/eda.qmd
part62

@@
   )
-characteristics <- data.frame(characteristics)
+
+# Convert tbl_summary object to data frame and assign column names
+characteristics <- as.data.frame(characteristics)
+column_names <- c("Characteristic", "No TB", "With TB")
+names(characteristics) <- column_names
+
 # Print the summary table
-characteristics
+print(characteristics)
 

### 2024-05-05 reviews/HardinParker_Kaggwa_project-review-template.qmd
part621

---
title: Project Review Template 
author: Emma Hardin-Parker
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING

Name of project author(s): Patrick Kaggwa

Name of project reviewer: Emma Hardin-Parker

# Instructions

# Specific project content evaluation

## Background, Context and Motivation

### Feedback and Comments

Hey Patrick, I would suggest revising your Introduction section of the Abstract. It is a little tough to understand on a first read. You provided a great deal of insight into the background and described the data and data collection well. Past research was not mentioned; however, you provided enough references and mentioned tools already used in Uganda so I think it is okay! I would suggest adding some information on why you are "motivated" to do this project and what you hope to gain from it.

### Summary assessment

* strong contextualization and motivation

## Question description

### Feedback and Comments

The questions and hypotheses are well defined. The questions relate well to the data and are very clear.

### Summary assessment

* question/hypotheses fully clear

## Data description

### Feedback and Comments

The data is described very well and the README.md file within the raw-data file contains a detailed codebook. I felt very comfortable understanding the data and how/why the data were collected. The source was also provided.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis

### Feedback and Comments

The data is well cleaned and explored; however, there were not many explanations on the "why" behind your processing and exploratory analysis. Adding some information in the eda.qmd or processingfile-v1.qmd file may be helpful! The Supplementary-Material.qmd file is untouched and unedited. I recommend following some of Dr. Handel's guidance and the Project Examples on the website to ensure this file contains the right material!

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

I would recommend adding more information in the README.md file within the analysis-code folder. Since you have multiple Quarto files with different analyses, explanation on why they are separate or where to go to reproduce the results would be helpful. I would continue to add more explanations within these files to explain why you are conducting this analysis and what each step is telling you. The analyses were all appropriate for the data and the analyses were conducted properly.

### Summary assessment

* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results are presented well and are easy to read and understand. If you have time, I would suggest making some improvements to plots and tables so they print more professionally. Some features in plotly() and ggplot2() can really help make your figures and tables look nicer! It can be a pain but it really pays off.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions

### Feedback and Comments

The findings are well reported in the Manuscript. Both strengths and limitations are acknowledged in depth and the findings are accurately reported.

### Summary assessment

* strong, complete and clear discussion

## Further comments

Patrick, you've done great work. I would really recommend explaining your processes more within the individual analysis and processing files. I would also recommend really bolstering both your README.md and Supplementary-Material files. The Supplementary-Material file is crucial to be able to formally present results and processes that were not included with the manuscript itself.

# Overall project content evaluation

## Structure

### Feedback and Comments

The project is well-structured, but some work needs to be done with the README.md files and Supplementary-Material file. Some folders are unusued/contain untouched README.md files so you may need to address those things. All files are named appropriately and all figures and tables are labeled perfectly.

### Summary assessment

* mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

## Documentation

### Feedback and Comments

The manuscript is documented very well, but some of that documentation is lost throughout the rest of the project. README.md files are crucial to having full understanding of how the analysis works and where everything is supposed to be. Additionally, I would recommend putting comments or introductions in your analysis files to ensure that any viewer knows exactly what is going on.

### Summary assessment

* decently documented with some gaps

## Reproducibility

### Feedback and Comments

All of your results were fully reproducible. There weren't a lot of files that clearly explained "how" to reproduce the results. I was able to just from our experience with this class and this sort of project, so you may need to bolster those explanations. I was able to run everything without manual interventions or edits.

### Summary assessment

* fully reproducible without issues

## Thoroughness

### Feedback and Comments

The project overall is very thorough. You really dove into a lot of ML models and techniques we learned over the semester! I really commend you for that. You did a great job discussing the data and the analyses themselves in the manuscript. Alternative methods were discussed in the limitations section and allow for further insight/questions into what kinds of studies/research could be conducted in the future. Overall, you did a great job!

### Summary assessment

* strong level of thorougness

## Further comments

Your project is great! I added some comments/suggestions throughout this file. Please do not hesistate to reach out if you have any clarifying questions. I really enjoyed reading your manuscript and going through your analyses!

### 2024-05-05 reviews/hall-project-review-k.qmd
part621

---
title: Project Review 
author: ELIZABETH HALL
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING

Name of project author: Patrick Kaggwa

Name of project reviewer: Elizabeth Hall

------------------------------------------------------------------------

# Specific project content evaluation

------------------------------------------------------------------------

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of the project is well described, and provides a comprehensive background with summaries of previous/related work, with multiple citations listed. The project itself fits well into the context of the existing body of work. The motivation for the project is also clear, as the need for faster diagnosis times is mentioned in the background and the study looks at early clinical TB data. What new information the study hopes to provide is not explicitly stated, but can be inferred based on the background provided.

### Summary assessment

-   strong contextualization and motivation

------------------------------------------------------------------------

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions/hypotheses are explicitly stated, and have a dedicated section in the manuscript. It is also clear how the questions relate to the data.

### Summary assessment

-   question/hypotheses fully clear

------------------------------------------------------------------------

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The source of the data is not given (that I could find), but the data collection is detailed in the manuscript. The data structure is somewhat explained in the `README.md` file in the `raw-data` folder, definitions are provided but specific variable names are not provided.

I would suggest making a simple codebook to explain each variable, and link back to the data source. GitHub has various codebook templates to chose from/reference. I used [this](https://gist.github.com/fmaskari/5886dbcc5ab73f2ea663) template for my project.

### Summary assessment

-   source and overall structure of data somewhat explained

------------------------------------------------------------------------

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

WRITE YOUR FEEDBACK HERE

The data is well cleaned and annotated. The data processing steps are neatly documented and explained in the `processingfile-v1.qmd` file which is in the `processing-code` folder. Alternatives for data wrangling/exploratory analysis are not explicitly discussed.

Exploratory code is also well cleaned and annotated, with steps being documented in `eda.qmd` which is in the `eda-code` folder. Alternatives are not discussed. Meaningful results are shown which can be seen when running the code or found in the `figures` folder under `results`.

I would recommend mentioning alternative exploratory/data wrangling methods, but otherwise this part looks good!

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

------------------------------------------------------------------------

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis methods used were appropriate for the data, and the analysis was done properly. Metrics like AUC and accuracy were used to evaluate model performance, which based on the brief searching I did, is appropriate for the data. The different components of the analysis are also done well, and explained thoroughly in the qmd files that accompany the code.

Your analysis section seems thorough and well done!

### Summary assessment

-   strong and reasonable analysis

------------------------------------------------------------------------

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results are presented decently, but there is room for improvement. Tables/figures have informative descriptions but no titles or labels so they are not publication quality.

I'd recommend changing variable names when graphing/generating tables so that they don't get cut off and are easier to read. I'd also recommend adding graph/table titles so that it's content can be easily contextualized. My last recommendation is to add figure/table labels (e.g. "Figure 1", "Table 1", etc.)

### Summary assessment

-   results are presented ok, with room for improvement

------------------------------------------------------------------------

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The findings are are discussed properly, and explained well. This section could be fleshed out with references to other literature. Strengths and limitations are explicitly acknowledged and the findings are interpreted properly.

### Summary assessment

-   strong, complete and clear discussion

------------------------------------------------------------------------

## Further comments

Your project looks good so far! My main suggestions are to do some work on your tables/figures, and to include more information about the data.

------------------------------------------------------------------------

# Overall project content evaluation

------------------------------------------------------------------------

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The structure of the project is mostly clear but could be improved. Some readme's still contain text from the original template, and should be updated or removed. Other than cleaning up junk files, the structure is good!

### Summary assessment

-   mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

------------------------------------------------------------------------

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project is well documented with corresponding qmd files for most code. These qmd files are well annotated, and the explanations are easy to follow. The code itself also contains comments which provide less in-depth, but clear explanations of the code's function.

### Summary assessment

-   fully and well documented

------------------------------------------------------------------------

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was able to reproduce most results with little to no issue, however I encountered errors when running the `statistical-analysis.R` code, which is in the `analysis-code` folder. Machine learning models can be finicky, but this is a fixable issue.

I also encountered minor issues with the `edacode.R` involving path issues when loading the data which I was able to fix via manual intervention.

### Summary assessment

-   small parts not reproducible or required manual intervention

------------------------------------------------------------------------

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Overall the study was decently thorough. The questions/hypotheses were fully addressed, though alternatives weren't really discussed. Discussion of alternative methods of data processing and different models could bolster the study in this area.

### Summary assessment

-   decent level of thoroughness

------------------------------------------------------------------------

## Further comments

Your project looks good so far! All issues I came across are easily fixable, and I especially liked your analysis section. My suggestions for this part are to make sure your code is reproducible and to mention some alternatives. Great job! :]

### 2024-05-05 reviews/project-review-template-ruiz.qmd
part621

---
title: Project Review Template 
author: Andrew Ruiz
date: 2024-04-25
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project:
CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING
Name of project author(s):
Patrick Kaggwa
Name of project reviewer:
Andrew Ruiz

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background, context and motivation were clear and to the point. However, this section could benefit from more details on the potential impact the work. Currently, the problem is that TB cases are not idenitified quickly enough and there are undiagnosed cases. Early detection could help reduce the spread of TB as well.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions are well defined, but could benefit from some refinement. Consider adding a narrower focus to question 1. "How can (TB) prediction be modeled using early clinical patient data from Kampala residents who visited the IDRC between 2011 and 2018, presenting with a persistent cough lasting two weeks or more?"

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
data is well described and the codebook is available. 
 The data table on page 4 is a little confusing to read. It could be improved by adding more descriptive headers and using a format that clearly defines the rows. The NAs in the columns do not seem necessary
### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The EDA seems reasonable and is well documented. The objectives of each step were clear. The coding and processes were uncluttered and easy to follow.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis seems appropriate and the metrics were clearly defined.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
Teh results are generally well presented. However, the figures should include title or captions. Because the graphs look very similar, it might be helpful to show them faceted or on the same figure for easier comparison.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Discussion and conclusions are appropriate. Strengths and limitations are acknowledged. The discussion could be improved by providing more context on the implications of the findings -could this lead to cost-savings, more efficient use of resources, better patient outcomes, etc.

This is not hugely important for this paper, but are current data collection methodolgies (paper vs electronic) in place a barrier for implementing models?

### Summary assessment
* strong, complete and clear discussion

## Further comments

Well done! I suggest reviewing the Word doc for formatting issues. Spacing for in-text citations was incorrect. Some errors in punctuation and capitalization were present.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

WRITE YOUR FEEDBACK HERE

### Summary assessment
* poor and confusing structure
* mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

WRITE YOUR FEEDBACK HERE

### Summary assessment
* poorly documented
* decently documented with some gaps
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The code ran smoothly with the exception of two errors. Both errors occurred in the analysis qmd. 
they are outlined below.
```{r}
#ML_analysis.qmd
#Line 242
#I modified the code to this and it worked: 
final_fit_rf <- finalize_workflow(rf_wf, select_best(tune_outputrf, metric = "roc_auc")) %>% 
  fit(data = train)
#Line 322
#I modified the code to this and it worked: 
svmfinal_fit <- finalize_workflow(svm_wf, select_best(tune_outputsvm, metric = "roc_auc")) %>% 
  fit(data = train)

```

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The study addressed the objectives outlined. The technical descriptions of the models and results were appropriate. THe conclusion could be improved by explicitly addressing the questions who this work addressed the questions posed in the introduction.

### Summary assessment
* strong level of thorougness

## Further comments

The biggest distraction for me was the data table. Consider ways to improve the table and make it easier to read.

### 2024-05-05 products/manuscript/Manuscript.qmd
part63

@@
 ## Introduction
-Although early detection and treatment of Tuberculosis cases are the hallmark of successful TB control, diagnostic delay is still long and common in Uganda.
-Therefore, the study aimed at developing a tool that uses machine-learning techniques to detect and diagnose TB more accurately in a shorter time.
+Although early detection and treatment of Tuberculosis cases are the hallmarks of successful TB control, diagnostic delays are still long and common in Uganda.
+Therefore, this study aimed to develop a tool that utilizes machine-learning techniques to detect and diagnose TB more accurately within a shorter time frame.
 
 ## Methods
-This was a retrospective study that used secondary data collected between 2011 to 2018 collected from health under Infectious diseases research collaboration an (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
-The study was conducted on data from 2296 tested patients of whom 1345 had no TB and 951 had TB.
-Different machine learning metrics(Accuracy, RMSE and AUC) were used evaluate the performance of the models.
+This was a retrospective study that utilized secondary data collected between 2011 and 2018 from health facilities under the Infectious Diseases Research Collaboration (IDRC) project called MIND (Mulago Inpatient Non-invasive Diagnosis for pneumonia).
+The study analyzed data from 2296 tested patients, among whom 1345 were found to have no TB and 951 were diagnosed with TB.
+Various machine learning metrics (Accuracy, RMSE, and AUC) were used to evaluate the performance of the models.
 
 ## Results
-Evaluation of four machine learning models—LASSO, Random Forest, Support Vector Machine (SVM), and Null model—revealed notable performance differences.
-Notably, the Random Forest model achieved the highest accuracy at 81.57% and an AUC of 0.8868, outperforming the other models.
+Evaluation of four machine learning models, LASSO, Random Forest, Support Vector Machine (SVM), and Null model—revealed notable performance differences.
+Particularly, the Random Forest model achieved the highest accuracy at 81.57% and an AUC of 0.8868, outperforming the other models.
 
-## Consclusion
-The results highlight the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies, offering valuable insights for clinical decision-making and the development of more effective screening protocols.
+## Conclusion
+The results underscore the potential of advanced machine learning algorithms, particularly Random Forest, in enhancing TB diagnosis and early detection strategies.
+These findings offer valuable insights for clinical decision-making and the development of more effective screening protocols.
+
 {{< pagebreak >}}
@@
 This delay impacts individual prognosis and community transmission.
-TB, caused by the bacterium Mycobacterium tuberculosis, primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@baykan_extrapulmonary_2022],.
+TB caused by the bacterium Mycobacterium tuberculosis primarily affects the lungs but can also involve other organs in its active form or remain latent without symptoms[@baykan_extrapulmonary_2022].
 In 2020, TB ranked among the top 10 causes of death globally, with over 10 million new cases and 1.3 million deaths reported.
-Africa bears a considerable burden, with an estimated quarter of new cases globally, resulting in approximately 417,000 deaths annually[@chakaya_who_2022].
+Africa bears a considerable burden with an estimated quarter of new cases globally, resulting in approximately 417,000 deaths annually[@chakaya_who_2022].
 TB is a leading cause of death among HIV-infected patients in sub-Saharan Africa, compounded by the low sensitivity of the commonly used sputum smear microscopy, particularly in detecting TB among people living with HIV[@palattiyil_access_2022].
@@
 ## Questions/Hypotheses to be addressed
-The prediction of TB can be modeled using early clinical patient data from Kampala residents who visited the IDRC between 2011 and 2018 and presented with a persistent cough lasting two weeks or more.
+How can the prediction of TB be modeled  early using clinical patient data from residents of Kampala who visited the IDRC between 2011 and 2018 who had cough for more than two weeks?
 
-{{< pagebreak >}}
 # Methods
@@
 Random Forest (RF): RF is a versatile and robust ensemble learning method that aims to reduce variance by building multiple decision trees on random subsets of the data and then averaging their predictions.
-It can handle large datasets with high-dimensional feature spaces and is less prone to overfitting compared to individual decision trees[@simon_interpreting_2023].
+It can handle large datasets with high-dimensional feature spaces and is less prone to over fitting compared to individual decision trees[@simon_interpreting_2023].
 However, RF models are more complex and may be challenging to interpret compared to simpler models like decision trees.
@@
 It works by finding the optimal hyperplane that separates data points belonging to different classes with the maximum margin of separation.
-SVM is effective in high-dimensional spaces and is versatile in handling linear and non-linear relationships through different kernel functions.[@javatpoint_support_2021] However, SVM can be computationally intensive, especially with large datasets, and may require careful tuning of hyperparameters for optimal performance.
+SVM is effective in high dimensional spaces and is versatile in handling linear and non-linear relationships through different kernel functions.[@javatpoint_support_2021] However, SVM can be computationally intensive, especially with large datasets, and may require careful tuning of hyperparameters for optimal performance.
 
@@
 #| label: tbl-schematic2
-#| tbl-cap: "Random Forest"
+#| tbl-cap: "Random Forest Predictions"
 rf <- readRDS(here("results", "tables", "rfconf_table.rds"))
@@
 #| label: tbl-schematic5
-#| tbl-cap: "Support Vector Machine"
+#| tbl-cap: "Support Vector Machine Predictions"
 svm<- readRDS(here("results", "tables", "svmconf_table.rds"))
@@
 #| label: tbl-schematic3
-#| tbl-cap: "lassoconf_table"
+#| tbl-cap: "LASSO Predictions"
 ls<- readRDS(here("results", "tables", "lassoconf_table.rds"))
@@
 
-# Models Evaluation
+## Models Evaluation
 
@@
 
-The study presents a comprehensive analysis of machine learning models' efficacy in predicting the presence of tuberculosis (TB) based on a dataset comprising 2296 instances and 14 variables associated with TB.
-The study selected nine variables for prediction after conducting chi-square tests to eliminate non-statistically significant variables associated with the outcome.
-The retained variables for prediction included asthma, alcohol use, fever, chest radiography, home fuel use, weight loss, presence of blood in sputum, oxygen saturation, and HIV status.
-The study evaluated four distinct machine learning models, namely LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, for their predictive performance using key metrics accuracy and Area Under the Curve (AUC).
-The LASSO model achieved an accuracy of 80.26% and an AUC of 0.8895.
-The Random Forest model outperformed others with an accuracy of 81.57% and an AUC of 0.8868.
+The study offers a comprehensive analysis of machine learning models' effectiveness in predicting tuberculosis (TB) presence using a dataset containing 2296 instances and 14 variables associated with TB.
+The study employed cross-validation with the LASSO method to identify significant predictors of the outcome.
+Three distinct machine learning models, LASSO, Random Forest, and Support Vector Machine (SVM) were evaluated for predictive performance using accuracy and Area Under the Curve (AUC) metrics.
+The LASSO model achieved an accuracy of 80.26% and an AUC of 0.8895, while the Random Forest model outperformed others with an accuracy of 81.57% and an AUC of 0.8868.
 SVM achieved an accuracy of 79.68% and an AUC of 0.8825.
-The Null model achieved an accuracy of 60.23% and an AUC of 0.5.
+In contrast, the Null model achieved an accuracy of 60.23% and an AUC of 0.5.
+Detailed classification results were presented for each model, highlighting their ability to correctly classify true positives and true negatives while minimizing false positives and false negatives.
+Notably, the Random Forest model demonstrated robust performance in accurately identifying true positives and true negatives with minimal misclassifications.
 
-Furthermore, the study presented detailed classification results for each model, emphasizing their ability to correctly classify true positives and true negatives while minimizing false positives and false negatives.
-Particularly, the Random Forest model demonstrated robust performance in correctly identifying true positives and true negatives, with minimal misclassifications.
-The findings underscore the potential of advanced machine learning algorithms, particularly Random Forest and LASSO, in enhancing TB diagnosis and early detection strategies.
-The manuscript provides valuable insights into the predictive capabilities of different machine learning models which can inform clinical decision-making and aid in the development of more effective TB screening tolls and protocols.
+## Strengths
 
-# Strengths
-
-In this study, a rigorous variable selection process was implemented, utilizing chi-square tests to identify statistically significant predictors associated with tuberculosis (TB) from a dataset comprising 2296 instances and 14 variables.
-By employing stringent selection criteria, the study ensures that only variables with meaningful associations with TB are included in subsequent analyses, thereby enhancing the model's interpretability, and reducing the risk of overfitting.
-Moreover, the evaluation of multiple machine learning models, including LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, contributes to a comprehensive comparison of different predictive approaches.
+Employing cross-validation helped enhance the model’s interpretability and reduce the risk of overfitting.
+Moreover, evaluating multiple machine learning models, including LASSO, Random Forest, Support Vector Machine (SVM), and a Null model, contributes to a comprehensive comparison of different predictive approaches.
 This assessment allows for informed decision-making regarding model selection, facilitating the identification of the most suitable algorithm for TB prediction.
-
+Additionally, the study employs robust performance metrics such as accuracy and Area Under the Curve (AUC) to evaluate model performance objectively.
+These metrics provide a comprehensive assessment of each model’s predictive capabilities, enabling comparisons across different algorithms.
+Furthermore, the detailed classification results presented offer insights into the models’ ability to correctly classify true positives and true negatives while minimizing false positives and false negatives.
 Additionally, the study employs robust performance metrics such as accuracy and Area Under the Curve (AUC) to evaluate model performance objectively.
@@
 
-# Limitations and Recommendations
+## Limitations and Recommendations
 
@@
 Firstly, the dataset's potential class imbalance, characterized by variations in the number of instances for individuals with and without TB, may impact model performance of the model.
-Additionally, while the variable selection methodology employed chi-square tests, alternative feature selection methods like recursive feature elimination could offer further insights into the predictive variables' relevance and importance.
-Future research should aim incorporate larger, more diverse datasets and conducting external validation studies to assess model performance in real-world clinical settings and use of more robust techniques such as neural networks.
+Future research should aim incorporate larger, more diverse datasets and conducting external validation studies to assess model performance in real world clinical settings and use of more robust techniques such as neural networks.
 
-# Conclusions
+## Conclusions
 

### 2024-05-05 products/supplement/supplement.qmd
part63

---
title: "CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING"
subtitle: ""
author: Patrick Kaggwa
date: "2024-02-24"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/MADAPROJECT.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(magrittr)
library(flextable)
library(htmltools)
```

## Supplemental Information
This document represents additional information about the data and results.

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic2
#| fig-cap: "TB_barplot"
#| echo: false
knitr::include_graphics(here("results","figures", "tb_barplot.png"))
```
```{r echo=FALSE, message = FALSE}
#| label: fig-schematic3
#| fig-cap: "Age distribution"
#| echo: false
knitr::include_graphics(here("results","figures", "age_distribution.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic10
#| fig-cap: "AGe _TB Box PLot"
#| echo: false
knitr::include_graphics(here("results","figures", "age_boxplot.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic4
#| fig-cap: "Hivstatus barplot"
#| echo: false
knitr::include_graphics(here("results","figures", "hivstatus_barplot.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: tbl-schematic8
#| tbl-cap: "Table showing Characteristics"
ls<- readRDS(here("results", "tables", "characteristics.rds"))
ls %>%
  knitr::kable()
```
